On First-order Proofs without syntax

### 2020-12-23 main.tex
Add some basic definitions

@@

\bfit{Terms} $s, t, u, \ldots$ and \bfit{atoms} (atomic formulas) $\alpha, \beta,
\gamma, \ldots$ are generated inductively from variables $x, y, z, \ldots$: if
$f$ (resp. $p$) is an $n$-ary function (resp. predicate) symbol and $t_1, \ldots,
t_n$ are terms, then $ft_1 \ldots t_n$ (resp. $pt_1 \ldots t_n$) is a term
(resp. atom). The set of atoms is extended with logical constant $\ttt$ and
$\fff$. We also assume that each predicate symbol $p$ has a \bfit{dual}
predicate symbol $\overline{p}$ with $\overline{p} \neq p$ and
$\overline{\overline{p}} = p$. We extend the duality to atoms with
$\overline{pt_1 \ldots t_n} = \overline{p}t_1 \ldots t_n$, $\overline{\ttt} =
\fff$, and $\overline{\fff} = \ttt$. \bfit{Formulas} $f, g, h$ are generated inductively from atoms by binary
connectives $\vee$ and $\wedge$ and quantifiers $\forall x$ and $\exists x$.
The \bfit{negation} $\neg$ is defined in the usual way.

A \bfit{directed graph} $G = (V, E)$ consists of a finite set $V$ of
\bfit{vertices} and a set of $E \subseteq V \times V$ of \bfit{direct edges}. An
\bfit{undirected graph} $G = (V, E)$ consists of a finite set $V$ of vertices
and a set of $E$ of \bfit{undirected edges} (two-element subsets of $V$).
For a given (directed or undirected) graph $G$, we write $V_G$ and $E_G$ for its
vertex and edge sets respectively. We write $vw$ for an edge $(v, w)$ or $\{v,
w\}$. Let $G = (V, E)$ and $G' = (V', E')$ be two undirected graphs. Without
loss of generality, we assume that $V$ and $V'$ are disjoint. The \bfit{union}
$G + G'$ is the graph $(V \cup V', E \cup E')$ and the \bfit{join} $G \times G'$
is the graph $(V \cup V', E \cup E' \cup \{vv' | v \ in V, v' \in V'\})$. A
\bfit{cograph} is a $P_4$-free undireced grpah, i.e., for any distinct vertices $v_1,
v_2, v_3, v_4 \in V$, the restriction of the edge sets on them is not equal to
$\{v_1v_2, v_2v_3, v_3v_4\}$. In a graph $G = (V, E)$, the \bfit{neighbourhood}
$N(v)$ of $v \in V$ is defined as the set $\{w | vw \in E\}$, and a module is a
set $M \subseteq V$ such that $N(v) \ M = N(w) \ M$ for all $v, w \in M$. A
\bfit{module} $M$ is strong if for every module $M'$, we have $M' \subseteq M$, $M
\subseteq$ or $M \cap M' = \emptyset$.

\subsection{First-order graphs}
A cograph is \bfit{logical} if every vertex is labelled either by an atom or variable,
and it has at least one atom-labelled vertex. An atom-labelled vertex is called
a \bfit{literal} and a variable-labelled vertex is called a \bfit{binder}. A
binder labelled with $x$ is called an \bfit{$x$-binder}. The \bfit{scope} of a binder
$b$ is the smallest strong module containing at least two vertices, including
the binder itself.
In a logical cograph $G$, a binder $b$ is \bfit{existential} (resp. \bfit{universal}) if, for every other
vertex $v$ in its scope, we have $bv \in E_G$ (resp. $bv \notin E_G$). An
$x$-binder is \bfit{legal} if its scope contains no other $x$-binder and at least one
literal.

@@

\input{bibliography}

\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

>>>>

% ** SPECIAL **

\newcommand*{\Gr}{\mathcal{G}}
\newcommand*{\LG}{\mathcal{L}}

@@

\bfit{Terms} $s, t, u, \ldots$ and \bfit{atoms} (atomic formulas) $\alpha, \beta,
\gamma, \ldots$ are generated inductively from variables $x, y, z, \ldots$: if
$f$ (resp. $p$) is an $n$-ary function (resp. predicate) symbol and $t_1, \ldots,
t_n$ are terms, then $ft_1 \ldots t_n$ (resp. $pt_1 \ldots t_n$) is a term
(resp. atom). The set of atoms is extended with logical constant $\ttt$ and
$\fff$. We also assume that each predicate symbol $p$ has a \bfit{dual}
predicate symbol $\overline{p}$ with $\overline{p} \neq p$ and
$\overline{\overline{p}} = p$. We extend the duality to atoms with
$\overline{pt_1 \ldots t_n} = \overline{p}t_1 \ldots t_n$, $\overline{\ttt} =
\fff$, and $\overline{\fff} = \ttt$. \bfit{Formulas} $A, B, C$ are generated inductively from atoms by binary
connectives $\vee$ and $\wedge$ and quantifiers $\forall x$ and $\exists x$.
The \bfit{negation} $\neg$ is defined in the usual way. A formula is \bfit{rectified} if all bound variables are distinct from one another and from all free variables. Without loss of generality, we assume that all formulas are rectifed throughtout the paper as every formula has a logically equivalent rectified form.

A \bfit{directed graph} $G = (V, E)$ consists of a finite set $V$ of
\bfit{vertices} and a set of $E \subseteq V \times V$ of \bfit{direct edges}. An
\bfit{undirected graph} $G = (V, E)$ consists of a finite set $V$ of vertices
and a set of $E$ of \bfit{undirected edges} (two-element subsets of $V$).
For a given (directed or undirected) graph $G$, we write $V_G$ and $E_G$ for its
vertex and edge sets respectively. We write $vw$ for an edge $(v, w)$ or $\{v,
w\}$. Let $G = (V, E)$ and $G' = (V', E')$ be two undirected graphs. Without
loss of generality, we assume that $V$ and $V'$ are disjoint. The \bfit{union}
$G + G'$ is the graph $(V \cup V', E \cup E')$ and the \bfit{join} $G \times G'$
is the graph $(V \cup V', E \cup E' \cup \{vv' | v \in V, v' \in V'\})$. A
\bfit{cograph} is a $P_4$-free undireced grpah, i.e., for any distinct vertices $v_1,
v_2, v_3, v_4 \in V$, the restriction of the edge sets on them is not equal to
$\{v_1v_2, v_2v_3, v_3v_4\}$. In a graph $G = (V, E)$, the \bfit{neighbourhood}
$N(v)$ of $v \in V$ is defined as the set $\{w | vw \in E\}$, a set $W \subseteq
V$ induces a \bfit{matching} if for all $w \in W$, $N(w) \cap E$ is a singleton, and a module is a
set $M \subseteq V$ such that $N(v) \ M = N(w) \ M$ for all $v, w \in M$. A
\bfit{module} $M$ is strong if for every module $M'$, we have $M' \subseteq M$, $M
\subseteq$ or $M \cap M' = \emptyset$. A \bfit{graph homomorphism} $h: (V, E)
\rightarrow (V', E')$ is a function $h : V \rightarrow V'$ such that $h(v)h(w)
\in E'$ whenever $vw \in E$.

\subsection{First-order Graphs}
A cograph is \bfit{logical} if every vertex is labelled either by an atom or variable,
and it has at least one atom-labelled vertex. We denote by $label_G(v)$, or simply
$label(v)$ the label of a vertex in a logical cograph $G$. An atom-labelled vertex is called
a \bfit{literal} and a variable-labelled vertex is called a \bfit{binder}. A
binder labelled with $x$ is called an \bfit{$x$-binder}. The \bfit{scope} of a binder
$b$ is the smallest strong module containing at least two vertices, including
the binder itself. An \bfit{$x$-literal} is a literal whose atom contains the variable $x$. An $x$-binder \bfit{binds} every $x$-literal in its scope.
In a logical cograph $G$, a binder $b$ is \bfit{existential} (resp. \bfit{universal}) if, for every other
vertex $v$ in its scope, we have $bv \in E_G$ (resp. $bv \notin E_G$). An
$x$-binder is \bfit{legal} if its scope contains no other $x$-binder and at least one
literal.

We now introduce first-order graphs which are used to represent first-order formulas.

\begin{definition}
A \bfit{first-order graph} or \bfit{fograph} is a logical cograph with legal binders.
\end{definition}

For a formula $A$, we can define its associated fograph $\Gr(A)$ inductively by: $\Gr(\alpha) = \cdot \alpha$ (a single-vertex graph whose vertex is labelled by $\alpha$); $\Gr(A \vee B) = \Gr(A) + \Gr(B)$ and $\Gr(A \wedge B) = \Gr(A) \times \Gr(B)$.
The \bfit{binding graph} of a forgraph $G$ is the directed graph $(V_G, \{(b,
l) | b \text{ binds } l\})$.

\subsection{Fonets}
In a fograph, two atoms are \bfit{pre-dual} if their predicate symbols are dual
(e.g. $pxy$ and $\overline{p}yz$) and two literals are \bfit{pre-dual} if their
labels (atoms) are pre-dual. A \bfit{linked fograph} is a coloured fograph such
that every colour, called a \bfit{link}, consists of two pre-dual literals, and
every literal is either $\ttt$-labelled or in a link.

Let $G$ be a linked fograph. The set of links can be seen as a unification problem
by identifying dual predicate symbols. A \bfit{dualizer} of $G$ is an assignment
unifying all the links of $G$. Since a first-order unification problem is either
solvable or has a most general unifier, we can define the notion of \bfit{most
general dualizer}. A \bfit{dependency} is a pair $(\cdot x, \cdot y)$ of an
existential binder $\cdot x$ and a universal binder $\cdot y$ such that the most
general dualizer assigns to $x$ a term containing $y$. A \bfit{leap} is either a
link or a dependency. The \bfit{leap graph} $\LG_G$ of $G$ is the undirected graph $(V_G,
L_G)$ where $L_G$ is the set of leaps of $G$. A set $W \subseteq V_G$ induces a
\bfit{bimatching} in $G$ if $W$ induces a matching in $G$ and in the leap graph
$\LG_G$ as well.

\begin{definition}
A \bfit{first-order net} or \bfit{fonet} is a linked fograph which has
dualizer but no induced bimatching.  
\end{definition}

\subsection{Skew bifibrations}

A graph homomorphism $h: (V, E) \rightarrow (V', E')$ is a \bfit{fibration} if
for all $v \in V$ and $wh(v) \in E'$, there exists a unique $\tilde{w}$ such
that $\tilde{w}v \in E$ and $h(\tilde{w}) = w$, and is a \bfit{skew fibration}
if for all $v \in V$ and $wh(v) \in E'$ there exists $\tilde{w}$ such that
$\tilde{w}v \in E$ and $h(\tilde{w})w \notin E'$. A \bfit{fograph homomorphism}
is a graph homomorphism between the underlying graphs of the fographs
considered. A fograph homomorphism $h : G \rightarrow G'$ is
\bfit{label-preserving} if forall $v \in V_G, label(v) = label(h(v))$, and is
\bfit{existential-preserving} if for all existental binder $b$ in $G$, the
vertex $h(b)$ is an existential binder in $G'$.

\begin{definition}
A \bfit{skew bifibration} between fographs is a label-preserving and
	existential-preserving fograph homomorphism whose restriction to the
	binding graphs is a fibration.
\end{definition}

@@

%\input{bibliography}

%\begin{thebibliography}{1}

%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

%\end{thebibliography}
